Statistics for HIV surveillance

### 2021-09-24 docs/24-09-21_new-cohort/new-cohort.Rmd
Finish presentation for new cohort

---
title: "Spatio-temporal statistics for HIV surveillance"
author: "Adam Howes"
date: "August 2021"
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7
    fig_height: 3
    includes:
      in_header: preamble.tex
  slidy_presentation:
    highlight: haddock
subtitle: StatML Kick-off Camp
bibliography: citations.bib
institute: Imperial College London
---

@@

* Welcome to the CDT!
* Congratulations for many of you on completing your previous studies under some non-ideal circumstances

@@

* Some work in progress of mine, hopefully will make the above more concrete
* Young women are an important population for HIV prevention

>>>>

---
title: "Statistics for HIV surveillance"
author: "Adam Howes"
date: "August 2021"
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7
    fig_height: 3
    includes:
      in_header: preamble.tex
  slidy_presentation:
    highlight: haddock
subtitle: StatML Kick-off Camp
bibliography: citations.bib
institute: Imperial College London
---

@@

* Welcome to the CDT!
* Congratulations for many of you on completing your previous studies under some non-ideal circumstances
* PhDs can be challenging, learning process
    * I hope I'm a better researcher than I was two years ago
* Important to find the right work environment, draw on each other

@@

* Some work in progress of mine, to hopefully make the above concrete!
* Young women are an important population for HIV prevention
    * 10% of the population and 25% of all new infections [@unaids2021update]
* Not enough resources to provide programs for all young women in high prevalence settings $\implies$ standardised method for accessing population at risk required

##

\begin{table}
\centering
\begin{tabularx}{\textwidth}{lll}
\toprule
Category & HIV related risk & HIV risk ratio \\ 
 \midrule
None & Not sexually active & 0 \\
Low & Sexually active, one cohabiting partner & 1 \\
High & Non-regular sexual partner(s) & 1.72 \\
Very high & Young women from key populations & 13 \\
\bottomrule
\end{tabularx}
\caption{HIV risk categories and risk ratios.}
\end{table}

## Goal

* Estimate proportion of young women in each risk category
  * Districts $i = 1, \ldots, n$ in 13 AGYW Global Fund priority countries
  * Survey $t = 1, \ldots, T$
  * Age $a \in \{\text{15--19}, \text{20--24}, \text{25--29}\}$
* e.g. in Lilongwe (Malawi), in 2015, how many AGYW aged 20-24 are cohabiting with a single partner?

## Spatial random effects

* Tried none, IID and the Besag model (most common structured spatial random effect model)

$$
b_i \, | \, \mathbf{b}_{-i} \sim \mathcal{N} \left(\frac{\sum_{j: j \sim i} b_j}{n_{\delta i}}, \frac{1}{n_{\delta i}}\right),
$$

## Temporal random effects

* Tried none, IID and an AR1 model (most common structured spatial random effect model)

\begin{align*}
a_1 &\sim \left( 0, \frac{1}{1 - \rho^2} \right), \\
a_t &= \rho a_{t - 1} + \epsilon_t, \quad t = 2, \ldots, T, 
\end{align*}

##

\begin{table}[htb]
\centering
\begin{tabularx}{\textwidth}{llllllllll}
Model ID & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\ 
\midrule
Spatial structure & S1 & S2 & S3 & S1 & S2 & S3 & S1 & S2 & S3 \\
\midrule
Temporal structure & T1 & T1 & T1 & T2 & T2 & T2 & T3 & T3 & T3 \\
\end{tabularx}
\caption{All of the models considered}
\label{tab:models}
\end{table}

##

```{r echo=FALSE}
knitr::include_graphics("../../archive/process_information-criteria/20210916-133107-9c7e6acd/all-dhs-rank-comparison.pdf")
```
